### 2023-11-02 cover-letter/cover-letter.tex
Initial outline of cover-letter

\documentclass{letter}

\signature{Jerome Kelleher}

\address{Big Data Institute\\University of Oxford\\UK}
\begin{document}

\begin{letter}{GENETICS}

\opening{Dear Editors,}

I am writing on behalf of my coauthors to submit our manuscript entitled
\emph{A general and efficient representation of ancestral recombination graphs}
for consideration for publication in GENETICS
as a ``Theoretical Population and Evolutionary Genetics''
Investigations article.

\closing{Sincerely,}

\end{letter}
\end{document}