### 2024-04-18 cover-letter/cover-letter-resubmit.tex
Update cover letter

\documentclass{letter}

\signature{Jerome Kelleher}

\address{Big Data Institute\\University of Oxford}
\begin{document}

\begin{letter}{GENETICS}

\opening{Dear Graham,}

I am writing on behalf of my coauthors to resubmit our 
manuscript entitled
\emph{A general and efficient representation of ancestral recombination graphs}
for consideration for publication as a Perspective.

We are deeply grateful for your detailed feedback, which has allowed us 
to identify our target audience and clarified the purpose of this article.
We are absolutely in agreement that it is more suited to the Perspective
format, and are very happy for it to be considered potentially appropriate
for publication in Genetics.

We have heavily revised the manuscript, substantially shortening it
and moving some more technical and descriptive sections to appendices.
We have attached a detailed point-by-point response in the 
\texttt{response-to-reviewers.pdf} file, along with a 
latex-diff of the differences between the current and previous submissions.

Given your major influence on the manuscript, we have included you in the
Acknowledgements, and thank you again for taking the time to give
such helpful, thoughtful and detailed feedback.

\closing{Sincerely,}

\end{letter}
\end{document}