### 2024-06-04 cover-letter/cover-letter-resubmit-2.tex
Add cover letter

\documentclass{letter}

\signature{Jerome Kelleher}

\address{Big Data Institute\\University of Oxford}
\begin{document}

\begin{letter}{GENETICS}

\opening{Dear Graham,}

I am writing on behalf of my coauthors to resubmit our 
manuscript entitled
\emph{A general and efficient representation of ancestral recombination
graphs}.

We are delighed that it is potentially suitable for publication in GENETICS,
and have endeavored to address the points you have raised.

We have attached a detailed point-by-point response in the 
\texttt{response-to-reviewers-2.pdf} file, along with a 
latex-diff of the differences between the current and previous submissions.

Thank you again for your careful and helpful input throughout this process.

\closing{Sincerely,}

\end{letter}
\end{document}